\subsection{System timer}

System timer is simple 24bit wide timer that can be used for generating
periodical interrupt. This can be useful when you want to write some kind of
OS.

\subsubsection{Function}

System timer are clocked at same frequency as CPU and this cannot be changed. It
is counting from zero to up, until compare match with specified top occur. Then
interrupt request is generated, counter is set to zero and counting are restarted.

Timer have two registers, one for controlling and one with actual counter value. Reading
from registers are possible, writing too but when write is performed at counter
value register, counter is reseted and counting is restarted from zero.

With control register you can set top value and enable or disable timer.

\subsubsection{Registers}

All registers are listed in table \ref{tab:systim_reg_map}. Register SYSTVR is 24
bits wide and  contains  actual value of counter. Register SYSTCR is 25 bit wide
and contains value of top constant and timer enable bit. Bit 23 down 0 are this
top constant, bit 24 is timer enable bit.

\begin{table}[h]
    \centering
    \begin{tabular}{|l|l|l|}
        \hline
        \textbf{Offset} & \textbf{Name} & \textbf{Purpose}            \\ \hline
        $+0$            & SYSTCR        & SysTim control register.    \\ \hline
        $+1$            & SYSTVR        & SysTim value register.      \\ \hline
    \end{tabular}
    \caption{System timer register map}
    \label{tab:systim_reg_map}
\end{table}

\subsubsection{Hacking}

Because system timer is simple peripheral there is also nothing much to modify. Interface
consist from bus interface and one interrupt request signal. When you wand add another
timer, just remember to connect intrq to the interrupt controller.

\begin{lstlisting}[language=VHDL, frame=single]
entity systim is
    generic(
        BASE_ADDRESS: unsigned(23 downto 0) := x"000000"
    );
    port(
        clk: in std_logic;
        res: in std_logic;
        address: in unsigned(23 downto 0);
        data_mosi: in unsigned(31 downto 0);
        data_miso: out unsigned(31 downto 0);
        WR: in std_logic;
        RD: in std_logic;
        ack: out std_logic;
        intrq: out std_logic
    );
end entity systim;
\end{lstlisting}
